Outline

### 2021-01-11 text/main.Rmd
minor edits after comments S. Villéger

**Authors:** Nina M. D. Schiettekatte^1,2,*^, Simon J. Brandl^3^, Jordan M. Casey^3^, Nicholas A. J. Graham^4^, Diego R. Barneche^5,6^, Deron E. Burkepile^7,8^, Jacob E. Allgeier^9^, Jesús E. Arias-Gonzaléz^10^, Graham J. Edgar^11^, Carlos E. L. Ferreira^12^, Sergio R. Floeter^13^, Alan M. Friedlander^14^, Alison L. Green^15^, Michel Kulbicki^2,16^, Yves Letourneur^2,17^, Osmar J. Luiz^18^, Alexandre Mercière^1,2^, Fabien Morat^1,2^, Katrina S. Munsterman^9^, Enrico L. Rezende^19^, Fabian A. Rodríguez‐Zaragoza^20^, Rick D. Stuart-Smith^11^, Laurent Vigliola^2,16^, Sébastien Villéger^21^, Valeriano Parravicini^1,2^

^1^PSL Université Paris: EPHE-UPVD-CNRS, USR 3278 CRIOBE, Université de Perpignan, Perpignan, France;    
^2^Laboratoire d'Excellence “CORAIL”, Perpignan, France;     
^3^The University of Texas at Austin, Marine Science Institute, Port Aransas, TX 78373, USA;      
^4^Lancaster Environment Centre, Lancaster University, Lancaster, LA1 4YQ, UK;      
^5^Australian Institute of Marine Science, Crawley, WA, Australia;      
^6^Oceans Institute, The University of Western Australia, Crawley, WA, Australia;      
^7^Department of Ecology, Evolution, and Marine Biology, University of California, Santa Barbara, CA, USA;      
^8^Marine Science Institute, University of California, Santa Barbara, CA, USA;      
^9^Department of Ecology and Evolutionary Biology, University of Michigan, Ann Arbor, MI, USA;      
^10^Laboratorio Ecología de Ecosistemas de Arrecifes Coralinos, Departamento Recursos del Mar, CINVESTAV Unidad Mérida, AP73 Cordemex, CP97310, Mérida, Yucatán, Mexico;      
^11^Institute for Marine and Antarctic Studies, University of Tasmania, Hobart, TAS, Australia;      
^12^Departamento de Biologia Marinha, Universidade Federal Fluminense (UFF), Niterói,  the state of Rio de Janeiro (RJ), Brazil;     
^13^Marine Macroecology and Biogeography Lab, Depto. de Ecologia e Zoologia, CCB, Universidade Federal de Santa Catarina, Florianopolis, Santa Catarina, Brazil;      
^14^Department of Biology, University of Hawaii, Honolulu, HI, USA;      
^15^Red Sea Research Center, King Abdullah University of Science and Technology, Thuwal 23955-6900, Saudi Arabia;
^16^Institut de Recherche pour le Développement, UMR UR-IRD-CNRS-IFREMER-UNC ENTROPIE, Nouméa, New Caledonia, France;     
^17^Université de la Nouvelle-Calédonie, UMR UR-IRD-CNRS-IFREMER-UNC ENTROPIE, Nouméa, New Caledonia, France;     
^18^Research Institute for the Environment and Livelihoods, Charles Darwin University, Darwin, Australia;     
^19^Departamento de Ecología, Center of Applied Ecology and Sustainability (CAPES), Facultad de Ciencias Biológicas, Pontificia Universidad Católica de Chile, Santiago 6513677, Chile;      
^20^Laboratorio de Ecosistemas Marinos y Acuicultura (LEMA), Departamento de 12 Ecología, CUCBA, Universidad de Guadalajara. México. Carr. Guadalajara-Nogales 13 km. 15.5, Las Agujas Nextipac, Zapopan, C.P. 45110, Jalisco, 14, Mexico;     
^21^MARBEC, Université de Montpellier, CNRS, IFREMER, IRD, Montpellier, France     
^*^Correspondance to NMDS (nina.schiettekatte@gmail.com).

**One Sentence Summary:** A global assessment of multiple ecosystem functions reveals critical trade-offs among functions in coral reefs.

**Abstract:**
Preserving the integrity of coral reefs – their functioning – is a critical challenge of the 21st century.
However, a quantitative assessment of multiple functions is needed to guide conservation efforts.
Through the integration of empirically-parametrized bioenergetic models and global community surveys, we quantify five key functions mediated by fishes.
We show that functions exhibit critical trade-offs driven by diverging community structures.
Further, functions are locally dominated by few species, but worldwide, the identity of dominant species varies a lot - 70% of the 1110 species in our dataset are locally dominant.
Our quantitative assessment reveals a challenge in global coral reef conservation: no reef fish community can simultaneously sustain high values across all functions and there are no global keystone species to target.

**Main Text:**
The flow of elements through biological communities fuels all life on Earth [@Welti2017]. Preserving these fluxes, often termed ecosystem functions, is critical for the integrity of ecosystems [@Welti2017].
For millennia, resources have been managed with an economic mindset to maximize desirable functions such as the production of biomass [@Weisser2017].
Sustaining multiple functions requires both high species richness and a variety of species assemblages across a landscape [@Zavaleta2010].
However, there can be trade-offs between functions, and efforts to maximize one function can negatively impact another [@Zavaleta2010].
An understanding of such trade-offs is required to make informed management decisions, but simultaneously quantifying multiple ecosystem functions is challenging.
Therefore, trade-offs between functions, their drivers, and functional vulnerability are poorly understood in many ecosystems [@Brandl2019front].

Coral reefs are among the most diverse and productive ecosystems on Earth and provide essential ecosystem services [@Teh2013].
Yet, their integrity of is threatened by a plethora of anthropogenic stressors [@Hughes2017].
Severe declines in habitat quality and fish biomass have brought coral reef functioning to the forefront of scientific discourse [@Bellwood2019; @Brandl2019front].
However, our capacity to evaluate reef functioning relies on static proxies of functions, such as live coral cover, standing stock biomass of reef fishes, or functional richness based on qualitative traits [e.g. @Cinner2020].
Conversely, we know comparatively little about elemental fluxes and their drivers (but see [@Allgeier2016]), which constitutes a severe limitation to management [@Brandl2019front].

Here, we integrate biogeochemistry and community ecology to advance our understanding of the elemental fluxes that underpin reef fish functioning.
Using empirically-collected species-specific data on basic organismal processes and Bayesian phylogenetic models, we parameterize individual-level bioenergetic models to estimate five key ecosystem functions: nitrogen (N) excretion, phosphorus (P) excretion, biomass production, herbivory, and piscivory.
We apply these bioenergetic models to all individuals across 9,118 reef fish communities in 585 sites worldwide (Extended Data Table 1) to:
(1) quantify community-level reef fish functions and their trade-offs,
(2) extract the community- and species-level effects on these functions, and
(3) gauge the vulnerability of reef fish functioning in the Anthropocene.

The five key ecosystem functions performed by fishes across the world's reefs exhibit high variability (Fig. 1).
Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and we indeed observed a strong relationship between fish biomass and all five functions (Extended Data Fig. 1a-e, Extended Data Fig. 2).
However, our analyses demonstrate that functions vary remarkably after accounting for biomass - functions of communities with similar biomass may differ with two orders of magnitude or more (i.e. biomass production, Extended Data Fig. 1a-f).
Thus, using biomass as a proxy for functioning can mask differences in critical community-level functions.
Further, we demonstrate strong trade-offs among the five functions, independent of biomass (Fig.1, Extended Data Fig. 1g).
For example, high herbivory rates or nitrogen excretion negatively correlate with rates of phosphorus excretion.
Consequently, for a given value of biomass, no reef can yield above average values across all five functions.
While many reefs may stand out as hotspots for one function, none can maximize functioning as a holistic, multifaceted concept (Fig. 1).

Community structure and species-specific traits clearly impact rates of functioning.
First, using community-level ecological predictors known to affect elemental fluxes [@Schiettekatte2020] (body size, trophic level, species richness, biomass, temperature, and age structure; Fig. 2), we show that correlations between functions are mediated by contrasting aspects of community structure (Fig. 2; Extended Data Table 2; Extended Data Fig. 2).
For example, phosphorus excretion and piscivory are higher in communities with large-bodied fishes that occupy high trophic levels (Extended Fig. 1g; See @Allgeier2014).
In contrast, biomass production is highest in communities dominated by small and/or immature fishes at low trophic levels, creating a trade-off between biomass production and phosphorus excretion.
Metabolic theory predicts that small-bodied individuals have higher mass-specific metabolic rates, leading to elevated consumption rates and disproportional contributions to functions that rely on rapid energetic turnover [@Barneche2018; @Morais2020glob; @Brandl2019science].
Conversely, fishes in early life stages or with a nutrient-poor diet are often limited by phosphorus [@Schiettekatte2020], resulting in low contributions to phosphorus excretion.
Thus, due to variations in organismal physiology and life-history traits [@Schiettekatte2020], fish community structure can significantly impact ecosystem-wide functioning [@Schramski2015].

Secondly, alongside features of community structure, functions may also be influenced by specific high-performing taxa.
Certain species tend to contribute more or less to a particular function, relative to their contribution to biomass (Fig. 3a; Extended Data Fig. 3,4).
Further, abundant species impact rates of functioning at the community level [@Topor2019].
We show that functions consistently hinge on a few dominant species (Fig. 3b). Specifically, more than 50% of a given function, on average, is upheld by only 12% of the species present within a local community.
However, the identity of these species varies dramatically (Fig. 3c).
While few high-performing taxa dominate functioning in each location, there are no species that are important across their entire range, and most species are locally important in at least one community.
Indeed, 70% of all species contributed disproportionally to a specific function in at least one ommunity.
Despite high species richness on coral reefs, researchers often report the existence of functionally-dominant “key species” [e.g. @Bellwood2006].
Our results reveal that while functional dominance is indeed prevalent, the identity of local, dominant species vary strongly across different locations.

@@

Overall, through coral reef fishes, we demonstrate that the variability in processes that govern the cycling of elements presents an unrecognized challenge for coral reef ecosystem functioning. Management strategies that call for the enhancement of coral reef functioning via an economic mindset, where higher functioning is better, are not feasible. Instead, conserving coral reef ecosystem functioning will require a more nuanced approach that considers processes that vary beyond the effect of standing stock biomass and are subject to variable, local trade-offs, drivers, and anthropogenic threats.

>>>>

**Authors:** Nina M. D. Schiettekatte^1,2,\*^, Simon J. Brandl^3^, Jordan M. Casey^3^, Nicholas A. J. Graham^4^, Diego R. Barneche^5,6^, Deron E. Burkepile^7,8^, Jacob E. Allgeier^9^, Jesús E. Arias-Gonzaléz^10^, Graham J. Edgar^11^, Carlos E. L. Ferreira^12^, Sergio R. Floeter^13^, Alan M. Friedlander^14^, Alison L. Green^15^, Michel Kulbicki^2,16^, Yves Letourneur^2,17^, Osmar J. Luiz^18^, Alexandre Mercière^1,2^, Fabien Morat^1,2^, Katrina S. Munsterman^9^, Enrico L. Rezende^19^, Fabian A. Rodríguez‐Zaragoza^20^, Rick D. Stuart-Smith^11^, Laurent Vigliola^2,16^, Sébastien Villéger^21^, Valeriano Parravicini^1,2^

**Affiliations:**      
^1^PSL Université Paris: EPHE-UPVD-CNRS, USR 3278 CRIOBE, Université de Perpignan, Perpignan, France;    
^2^Laboratoire d'Excellence “CORAIL”, Perpignan, France;     
^3^The University of Texas at Austin, Marine Science Institute, Port Aransas, TX 78373, USA;      
^4^Lancaster Environment Centre, Lancaster University, Lancaster, LA1 4YQ, UK;      
^5^Australian Institute of Marine Science, Crawley, WA, Australia;      
^6^Oceans Institute, The University of Western Australia, Crawley, WA, Australia;      
^7^Department of Ecology, Evolution, and Marine Biology, University of California, Santa Barbara, CA, USA;      
^8^Marine Science Institute, University of California, Santa Barbara, CA, USA;      
^9^Department of Ecology and Evolutionary Biology, University of Michigan, Ann Arbor, MI, USA;      
^10^Laboratorio Ecología de Ecosistemas de Arrecifes Coralinos, Departamento Recursos del Mar, CINVESTAV Unidad Mérida, AP73 Cordemex, CP97310, Mérida, Yucatán, Mexico;      
^11^Institute for Marine and Antarctic Studies, University of Tasmania, Hobart, TAS, Australia;      
^12^Departamento de Biologia Marinha, Universidade Federal Fluminense (UFF), Niterói,  the state of Rio de Janeiro (RJ), Brazil;     
^13^Marine Macroecology and Biogeography Lab, Depto. de Ecologia e Zoologia, CCB, Universidade Federal de Santa Catarina, Florianopolis, Santa Catarina, Brazil;      
^14^Department of Biology, University of Hawaii, Honolulu, HI, USA;      
^15^Red Sea Research Center, King Abdullah University of Science and Technology, Thuwal 23955-6900, Saudi Arabia;
^16^Institut de Recherche pour le Développement, UMR UR-IRD-CNRS-IFREMER-UNC ENTROPIE, Nouméa, New Caledonia, France;     
^17^Université de la Nouvelle-Calédonie, UMR UR-IRD-CNRS-IFREMER-UNC ENTROPIE, Nouméa, New Caledonia, France;     
^18^Research Institute for the Environment and Livelihoods, Charles Darwin University, Darwin, Australia;     
^19^Departamento de Ecología, Center of Applied Ecology and Sustainability (CAPES), Facultad de Ciencias Biológicas, Pontificia Universidad Católica de Chile, Santiago 6513677, Chile;      
^20^Laboratorio de Ecosistemas Marinos y Acuicultura (LEMA), Departamento de 12 Ecología, CUCBA, Universidad de Guadalajara. México. Carr. Guadalajara-Nogales 13 km. 15.5, Las Agujas Nextipac, Zapopan, C.P. 45110, Jalisco, 14, Mexico;     
^21^MARBEC, Université de Montpellier, CNRS, IFREMER, IRD, Montpellier, France     
^\*^Correspondance to NMDS (nina.schiettekatte@gmail.com).

**One Sentence Summary:** A global assessment reveals critical trade-offs among multiple ecosystem functions performed by coral reef fishes.

\newpage

**Abstract:**
Preserving the functioning of coral reefs is a critical challenge of the 21st century.
However, a quantitative assessment of multiple functions is needed to guide conservation efforts.
Through the integration of empirically-parametrized bioenergetic models and global community surveys, we quantify five key functions mediated by fishes.
We show that functions exhibit critical trade-offs driven by diverging community structures.
Further, functions are locally dominated by few species, but worldwide, the identity of dominant species varies a lot - 70% of the 1110 species in our dataset are locally dominant.
Our quantitative assessment reveals a challenge in global coral reef conservation: no reef fish community can simultaneously sustain high values across all functions and there are no global keystone species to target.

**Main Text:**
The flow of elements through biological communities fuels all life on Earth [@Welti2017]. Preserving these fluxes, often termed ecosystem functions, is critical for the integrity of ecosystems [@Welti2017].
For millennia, resources have been managed with an economic mindset to maximize desirable functions such as the production of biomass [@Weisser2017].
Sustaining multiple functions requires both high species richness and a variety of species assemblages across a landscape [@Zavaleta2010].
However, there can be trade-offs between functions, and efforts to maximize one function can negatively impact another [@Zavaleta2010].
An understanding of such trade-offs is required to make informed management decisions, but simultaneously quantifying multiple ecosystem functions is challenging.
Therefore, trade-offs between functions, their drivers, and their vulnerability are still poorly understood in many ecosystems [@Brandl2019front].

Coral reefs are among the most diverse and productive ecosystems on Earth and provide essential ecosystem services [@Teh2013].
Yet, their integrity is threatened by a plethora of anthropogenic stressors [@Hughes2017].
Severe declines in habitat quality and fish biomass have brought coral reef functioning to the forefront of scientific discourse [@Bellwood2019; @Brandl2019front].
However, our capacity to evaluate reef functioning relies on static proxies of functions, such as live coral cover, standing stock biomass of reef fishes, or diversity of qualitative traits [e.g. @Cinner2020].
Conversely, we know comparatively little about actual functions - fluxes of elements and energy - and their drivers (but see [@Allgeier2016]), which constitutes a severe limitation to management [@Brandl2019front].

Here, we integrate biogeochemistry and community ecology to advance our understanding of the elemental fluxes that underpin reef fish functioning.
Using empirically-collected species-specific data on basic organismal processes and Bayesian phylogenetic models, we parameterize individual-level bioenergetic models to estimate five key ecosystem functions: nitrogen (N) excretion, phosphorus (P) excretion, biomass production, herbivory, and piscivory.
We apply these bioenergetic models to 9,118 reef fish communities in 585 sites worldwide (Table S1) to:
(1) quantify community-level reef fish functions and their trade-offs,
(2) extract the community- and species-level effects on these functions, and
(3) gauge the vulnerability of reef fish functioning in the Anthropocene.

The five key ecosystem functions performed by fishes across the world's reefs exhibit high variability (Fig. 1).
Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and we indeed observed a strong relationship between fish biomass and all five functions (Fig. S1a-e, Fig. S2).
However, our analyses demonstrate that functions vary remarkably after accounting for biomass - functions of communities with similar biomass may differ with two orders of magnitude or more (i.e. biomass production, Fig. S1a-f).
Thus, using biomass as a proxy for functioning can mask differences in critical community-level functions.
Further, we demonstrate strong trade-offs among the five functions, independent of biomass (Fig.1, Fig. S1g).
For example, high herbivory rates or nitrogen excretion negatively correlate with rates of phosphorus excretion.
Consequently, for a given value of biomass, no reef can yield above average values across all five functions.
While many reefs may stand out as hotspots for one function, none can maximize functioning as a holistic, multifaceted concept (Fig. 1).

Community structure and species-specific traits clearly impact rates of functioning.
First, using community-level ecological predictors known to affect elemental fluxes [@Schiettekatte2020] (body size, trophic level, species richness, biomass, temperature, and age structure; Fig. 2), we show that correlations between functions are mediated by contrasting aspects of community structure (Fig. 2; Table S2; Fig. S2).
For example, phosphorus excretion and piscivory are higher in communities with large-bodied fishes that occupy high trophic levels (Fig. S1g; See @Allgeier2014).
In contrast, biomass production is highest in communities dominated by small and/or immature fishes at low trophic levels, creating a trade-off between biomass production and phosphorus excretion.
Metabolic theory predicts that small-bodied individuals have higher mass-specific metabolic rates, leading to elevated consumption rates and disproportional contributions to functions that rely on rapid energetic turnover [@Barneche2018; @Morais2020glob; @Brandl2019science].
Conversely, fishes in early life stages or with a nutrient-poor diet are often limited by phosphorus [@Schiettekatte2020], resulting in low contributions to phosphorus excretion.
Thus, due to variations in organismal physiology and life-history traits [@Schiettekatte2020], fish community structure significantly impacts ecosystem-wide functioning [@Schramski2015].

Secondly, alongside features of community structure, functions may also be influenced by specific high-performing taxa.
Certain species tend to contribute more or less to a particular function, relative to their contribution to biomass (Fig. 3a; Fig. S3,S4).
Further, abundant species impact rates of functioning at the community level [@Topor2019].
We show that functions consistently hinge on a few dominant species (Fig. 3b). Specifically, more than 50% of a given function, on average, is upheld by only 12% of the species present within a local community.
However, the identity of these species varies dramatically between reefs (Fig. 3c).
While few high-performing taxa dominate functioning in each location, there are no species that are dominant across their entire range.
In addition, 70% of all species contributed disproportionally to a specific function in at least one community.
Despite high species richness on coral reefs, researchers often report the existence of functionally-dominant “key species” [e.g. @Bellwood2006].
Our results reveal that while functional dominance is indeed prevalent, the identity of local, dominant species vary strongly across different locations.

@@

Overall, through coral reef fishes, we demonstrate that the variability in processes that govern the cycling of elements presents an unrecognized challenge for protecting ecosystem functioning. Management strategies that call for the enhancement of coral reef functioning via an economic mindset, where higher functioning is better, are not feasible. Instead, conserving coral reef ecosystem functioning will require a more nuanced approach that considers processes that vary beyond the effect of standing stock biomass and are subject to variable, local trade-offs, drivers, and anthropogenic threats.
